---
title: "RovingDiverNutrient"
author: "Ashley"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}

if (!require("pacman")) install.packages("pacman") 
pacman::p_load(flextable, officer, stringr, ggplot2, stringr, gridExtra, ggpubr, Rmisc, FSA, rcompanion, RColorBrewer, dplyr, vegan, nparcomp, RVAideMemoire, MANOVA.RM, pairwiseAdonis, PMCMR, PMCMRplus, patchwork, plyr, gdata,paletteer, RColorBrewer, devtools, ochRe)
pacman::p_load_gh("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
```

```{r, load, include = "true"}
PD <- read.xls("NutrientRovingDiver.xlsx")
PD
PD$Date <- ymd(PD$Date)
PD$Date <- as.factor(PD$Date)
```

```{r}
PD$TotalObservations <- PD$SCTLD + PD$NoDisease
PD
```

```{r}
PD$NutrientAmmended = if_else(PD$Group %in% c("Experimental"), "Ammended", if_else(PD$Group %in% c("Healthy"), "Unammended", if_else(PD$Group %in% c("Control"), "Unammended","")))
PD

```

```{r}
#SpeciesData <- PD %>% group_by(Date) %>%
  

SpeciesData <- PD  %>% group_by(Date, SpeciesCode) %>%
 summarise_if(is.integer, sum)
 
 SpeciesPrevalenceData <-SpeciesData %>% group_by(Date) %>% mutate(SpeciesPrevalence = (TotalObservations)/ sum(TotalObservations)*100) %>%
   mutate(SCTLDSpecies = (SCTLD/sum(TotalObservations)*100)) %>% 
   mutate(SCTLDOnlySpecies = (SCTLD/TotalObservations) * 100) %>% arrange( desc(SCTLDOnlySpecies))

SpeciesPrevalenceData
```
```{r}
SpeciesPrevalenceSummary <- SpeciesPrevalenceData %>% group_by(SpeciesCode) %>%
  summarise(SpeciesMean = mean(SpeciesPrevalence), SCTLDMean= mean(SCTLDSpecies)) %>% arrange(desc(SpeciesMean))
SpeciesPrevalenceSummary

```


```{r}
TotalPrevalence <- PD %>% dplyr::select(c(Date, CattleTag, Group, SCTLD, NoDisease, TotalObservations, Status, NutrientAmmended, Temp, DiseaseStatus)) %>% group_by(Date,CattleTag, Group, Status, NutrientAmmended, Temp, DiseaseStatus) %>%
  summarise_if(is.integer, sum)
TotalPrevalence
```


#ADD A STATUS COLUMN INSTEAD OF MAKING A SEPERATE SPREAD SHEET THEN YOU AREN'T CONDENSING THE DATA
```{r}
StatusGrouped <- PD %>% dplyr::select(c(Status, SCTLD, NoDisease, TotalObservations)) %>%
  group_by(Status)
StatusGrouped
```
```{r}
DiseaseStatusGrouped <- PD %>% 

  dplyr::select(c(DiseaseStatus, SCTLD, NoDisease, TotalObservations, Temp, CattleTag)) %>%
    filter(!CattleTag %in% c("521","511")) %>%
  group_by(DiseaseStatus)
DiseaseStatusGrouped
```

```{r}
StatusGrouped$HealthyPrevalence = (StatusGrouped$NoDisease/StatusGrouped$TotalObservations * 100)
StatusGrouped$DiseasedPrevalence = (StatusGrouped$SCTLD/StatusGrouped$TotalObservations * 100)
StatusGrouped
```
```{r}
DiseaseStatusGrouped$HealthyPrevalence = (DiseaseStatusGrouped$NoDisease/DiseaseStatusGrouped$TotalObservations * 100)
DiseaseStatusGrouped$DiseasedPrevalence = (DiseaseStatusGrouped$SCTLD/DiseaseStatusGrouped$TotalObservations * 100)
DiseaseStatusGrouped
```

```{r}
TotalPrevalence$HealthyPrevalence = (TotalPrevalence$NoDisease/TotalPrevalence$TotalObservations *100)
TotalPrevalence$SCTLDPrevalence = (TotalPrevalence$SCTLD/TotalPrevalence$TotalObservations * 100)
```

# Looking at SCTLD Prevalence differences over group and time


```{r}

descdist(TotalPrevalence$SCTLDPrevalence)
```

```{r}
TotalPrevalence %>%
group_by(Group,Date) %>%
  shapiro_test(SCTLDPrevalence)

TotalPrevalence$SCTLDPrevalence_Log <- log(TotalPrevalence$SCTLDPrevalence + 1)

TotalPrevalence %>%
group_by(Group,Date) %>%
  shapiro_test(SCTLDPrevalence_Log)
```
```{r}

TotalPrevalence$CattleTag <- as.factor(TotalPrevalence$CattleTag)
TotalPrevalence$Date <- as.factor(TotalPrevalence$Date)
TotalPrevalence$Group <- as.factor(TotalPrevalence$Group)
TotalPrevalence$DiseaseStatus<- as.factor(TotalPrevalence$DiseaseStatus)


TotalPrevalence
```
#Running a mixed ANOVA to look at disease prevalence across time and groups
*Significant difference in timeand disease status* 
Mixed ANOVA (F (88,8) = 2.094, p = 0.044
```{r}
SCTLDPrevalence.ANOVA <-  #ungroup(TotalPrevalence) %>% 
  anova_test(data = as.data.frame(TotalPrevalence), dv = SCTLDPrevalence, wid = CattleTag, within= Date,between =Group)

get_anova_table(SCTLDPrevalence.ANOVA)
```

#Post-hoc tests looking at SCTLD Prevalence overtime 
```{r}



SCTLDPrevalence.pwc <- TotalPrevalence %>%
  group_by(Group) %>%
  pairwise_t_test(SCTLDPrevalence ~ Date,paired=TRUE, p.adjust.method = "bonferroni") %>%
  select(-df, -statistic, -p) # Remove details
SCTLDPrevalence.pwc
```



```{r}
TotalPrevalence %>%
  group_by(Date) %>%
  levene_test(SCTLDPrevalence ~ Status)
```



#Mixed ANOVA looking at if there is a significant difference in SCTLD prevalence around SCTLD or Healthy Colonies

*Significant difference in time* 
*Mixed ANOVA: F(8,104) = 2.148, p = 0.038

```{r}
Status.ANOVA <- 
  anova_test(data = as.data.frame(TotalPrevalence), dv = SCTLDPrevalence, wid = CattleTag, within= Date,between =Status)

get_anova_table(Status.ANOVA)
```
```{r}

SCTLDSTATUS.pwc <- TotalPrevalence %>%
  group_by(Status) %>%
  pairwise_t_test(SCTLDPrevalence ~ Date,paired=TRUE, p.adjust.method = "bonferroni") %>%
  select(-df, -statistic, -p) # Remove details
SCTLDSTATUS.pwc
```
#Going to look at disease status to see if there is any difference between active and inactive

Need to filter out colonies *511* & *521* with no status data and dates

```{r}
 ActiveStatusData <- TotalPrevalence %>% filter(CattleTag != "521" & CattleTag != "511") %>%
 filter(Date != "2021-09-27" & Date != "2021-10-05"& Date != "2021-10-15") 
ActiveStatusData
```

Mixed ANOVA to look at the effect of active status on SCTLD prevalence
***
  anova_test(data = as.data.frame(TotalPrevalence), dv = SCTLDPrevalence, wid = CattleTag, within= Date,between =Status)

get_anova_table(Status.ANOVA)
```{r}
ActiveStatusData <- TotalPrevalence %>% filter(!is.na(Temp)) %>% filter(DiseaseStatus != "")




```

AreaProportionStatus.Summary <- summarySE(ModelRateData, measurevar = "AreaProportion", groupvars = c("Date", "DiseaseStatus"), na.rm=TRUE) 
                                          
AreaProportionStatus.Summary
```{r}
SCTLDPrevalence.Summary <- summarySE(TotalPrevalence, measurevar = "SCTLDPrevalence", groupvars = c("Group", "Date"), na.rm= TRUE)
  
SCTLDPrevalence.Summary
```

Here we want to look at SCTLD prevelance over time for each group
```{r}
ggplot(Rate.Summary, aes(Timepoint,Rate, color=Group, group=Group))+
  geom_line()+
  geom_errorbar(aes(ymin= Rate-sd, ymax=Rate+sd), width=.5) 
```
```{r}
SCTLDPrevalence.Summary$Data <- as.factor(SCTLDPrevalence.Summary$Date)
SCTLDPrevalence.Summary$Group <- as.factor(SCTLDPrevalence.Summary$Group)
SCTLDPrevalence.Summary$Group <- factor(SCTLDPrevalence.Summary$Group,levels=c("Experimental", "Control", "Healthy"))
SCTLDPrevalence.Summary
```

***
I can keep this graph and group by month? 
```{r}
SCTLDPrevalence.LineGraph <- ggplot(SCTLDPrevalence.Summary,aes(Date,SCTLDPrevalence,color=Group,group=Group)) +
  geom_line(aes(color=Group), size=1.5)+
  geom_vline(xintercept= as.factor("2021-10-22"), linetype=4, color="black")+
#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("2021-10-22"), label="End Treatment", y=.17), colour="black",angle=90, vjust=2.5)+
  #adds label to intercept line
  geom_errorbar(aes(ymin=SCTLDPrevalence-se, ymax=SCTLDPrevalence+se),size=1.5, width=0.5) +
 scale_color_manual(values=c('#661100',"#999933", "#6699CC"), labels=c("Nutrient Amended", "Control", "Healthy")) +
  #adds colors to order of variables


  #scale_color_viridis_c()+
  #scale_fill_viridis_c() +
  #scale_color_manual(values=c('red2','lightcoral')) +
 scale_x_discrete(limits=c("2021-09-22","2021-09-27", "2021-09-30","2021-10-05", "2021-10-10","2021-10-15","2021-10-22", "2021-11-03", "2021-12-01"), labels=c("Sept 22","Sept 27","Sept 30","Oct 5", "Oct 10","Oct 15", "Oct 22", "Nov 11", "Dec 1")) +
  scale_y_continuous(breaks=c(0,5,10,15,20), labels=c("0%","5%", "10%", "15%", "20%"))  +
      labs(y="Surrounding SCTLD Prevalence", color="Experimental Group") +
  #scale_fill_discrete("Experimental Group") +
 #Change the x axis labels
  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"), 
        legend.key=element_blank())
      

SCTLDPrevalence.LineGraph
ggsave("Figures/SCTLDPrevalence.LineGraph.png", plot = SCTLDPrevalence.LineGraph, height = 5, width = 9, units = 'in', dpi = 600)
```



```{r}
Status.Summary <- TotalPrevalence %>%
  group_by(Status) %>%
  summarise(sd=sd(SCTLDPrevalence), Prevalence=mean(SCTLDPrevalence))
Status.Summary
```
***
Trying to visualize the differneces between just the diseased and apparently healthy colonies and surrounding disease prevalence 

```{r}
ggplot(TotalPrevalence,aes(Status,SCTLDPrevalence, color=Status)) +
  geom_boxplot()
```
***
Visualizing the SCTLD prevalence around the different colonies


```{r}
ggplot(TotalPrevalence,aes(Group,SCTLDPrevalence, color=Group)) +
  geom_boxplot()
```
```{r}
TotalPrevalence$Group <- as.factor(TotalPrevalence$Group)
TotalPrevalence$Data <- as.factor(TotalPrevalence$Date)

TotalPrevalence$Group <- factor(TotalPrevalence$Group,levels=c("Experimental", "Control", "Healthy"))
TotalPrevalence
```


```{r}
SCTLDPrevalence.Boxplot <- ggplot(data=TotalPrevalence, aes(Group,SCTLDPrevalence, color=Group)) +
  geom_violin() +
  geom_jitter(position = position_jitter(width = .2, height=0.07), alpha = 0.6) +#seed = 1, height= 0.055, width = 0.45)) +
  stat_summary(fun.y=median, geom="crossbar", width=.6, alpha=.8) +

 # geom_boxplot(aes(Group,SCTLDPrevalence, color=Group)) +
  scale_color_manual(values=c('#661100',"#999933", "#6699CC")) +
   scale_x_discrete(labels=c("Nutrient \nAmended", "Control", "Healthy")) +
  scale_y_continuous(breaks=c(0,5,10,15,20, 25), labels=c("0%","5%", "10%", "15%", "20%", "25%")) +
  
   labs(x="Experimental Group", y="Surrounding SCTLD Prevalence") +
  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"), 
        legend.position = "none")
SCTLDPrevalence.Boxplot
ggsave("Figures/SCTLDPrevalence.Boxplot.png", plot = SCTLDPrevalence.Boxplot, height = 5, width = 9, units = 'in', dpi = 600)
```

#Going to look at disease status to see if there is any difference between active and inactive
```{r}
anova_test(data = as.data.frame(ActiveStatusData), SCTLDPrevalence ~ Date*DiseaseStatus)
```

 #Effect of Temp on Overall Disease Prevalence
 
```{r}
ActiveStatusData

cor.test(ActiveStatusData$SCTLDPrevalence, ActiveStatusData$Temp, method="pearson")
?cor.test
```

 
 