---
title: "WaterQuality"
author: "Ashley"
date: "2022-08-25"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Basic setup of R environment

------------------------------------------------------------------------

## Loading required packages

For the following analyses we will require the use of a number of
different R packages. Most of which can be sourced from CRAN, but some
must be downloaded from GitHub. We can use the following code to load in
the packages and install any packages not previously installed in the R
console.

```{r, packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
```

```{r, packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman") 
pacman::p_load(flextable, officer, stringr, ggplot2, stringr, gridExtra, ggpubr, Rmisc, FSA, rcompanion, RColorBrewer, dplyr, vegan, nparcomp, RVAideMemoire, MANOVA.RM, pairwiseAdonis, PMCMR, PMCMRplus, patchwork, plyr, gdata,paletteer, RColorBrewer, devtools, ochRe, rstatix, lmer4, lmerTest, WRS2, Rstatix)

pacman::p_load_gh("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
```

##Uploading 3D Model Data from Excel *NutrientData* is Water data
organized into less rows and longer columns to make graphs easier

```{r, load, include = "true"}
WaterData <- read.xls("WaterQualityDataOnly.xlsx")
OrderedNutrientData<-read.xls("NutrientData.xlsx")
RainData <- read.xls("RainData.xlsx")
NewWaterData <- read.xls("NewNutrientData.xlsx")
NewWaterDataDIN <- read.xls("NewNutrientDataHorizontal.xlsx")
head(WaterData)
RainData
```

#Filtering Data

*GroupWaterData* = The data that contains all groups (*Healthy*,
*Experimental*, and *Control*) include except *TP4*, *TP6* and *TP8*

*HealthyWaterData* = The data of all timepoints filtering out the
*Control* and *Experimental* colonies since they were not measured at
all timepoints, this will allow us to look at differences of the ambient
groups *AmbientTreatmentData* = The healhty colonies during the
experiment

```{r}
GroupWaterData <- filter(WaterData, !Timepoint %in% c("TP4", "TP5", "TP6", "TP8"))
HealthyWaterData <- filter(WaterData, Group == "Healthy")
TreatmentWaterData <- filter(WaterData, Group == "Experimental") 
Exp.TreatmentWaterData <- filter(TreatmentWaterData, !Timepoint %in% c("TP0", "TP9"))
AmbientWaterData <- filter(WaterData, Treatment==  "No")
AmbientTreatmentData <- filter(WaterData, Treatment=="Yes", Group=="Healthy")
TreatmentWaterData
Exp.TreatmentWaterData
```

```{r}
Exp.TreatmentWaterData %>% summarise(NH4=mean(NH4), PO4 = mean(PO4), NO23 = mean(NO23))
```

```{r}
HealthyWaterData %>% summarise(NH4=mean(NH4), PO4 = mean(PO4), NO23 = mean(NO23))
```

```{r}
HealthyWaterData %>% get_summary_stats(NH4,NO23,DIN.SRP, show=c("min", "max", "mean"))
```

```{r}
Exp.TreatmentWaterData %>% get_summary_stats(NH4,NO23, show=c("min", "max", "mean"))
```

#Assessing Normality \*\*\* Using a Shapiro-Wilk Normality Test to test
normaility of one variable, we cannot assume normality because many of
the p values are \<.05

```{r}
WaterData %>%
group_by(Group,Timepoint) %>%
  shapiro_test(NH4)
```

Try log transforming the data to see if it makes it anymore normal

```{r}
WaterData$NH4_log<- log(WaterData$NH4)
WaterData
```

Test to see if log transformed data is anymore normal, and it is not

```{r}
WaterData %>%
group_by(Group,Timepoint) %>%
  shapiro_test(NH4_log)
```

#Viewing summary statistics

```{r}
AmbientTreatmentData %>% get_summary_stats(NH4, show=c("min", "max", "mean"))
```

```{r}
AmbientTreatmentData %>% get_summary_stats(NO23, show=c("min", "max", "mean"))
```

```{r}
AmbientTreatmentData %>%
  get_summary_stats(DIN, show=c("min", "max", "mean"))
```

```{r}
AmbientTreatmentData %>%
  get_summary_stats(PO4, show=c("min", "max", "mean"))
```

***Ambient water summary stats throughout whole experiment***

```{r}
AmbientWaterData %>%
  get_summary_stats(NO23, show=c("min", "max", "mean"))
```

```{r}
AmbientWaterData %>%
  get_summary_stats(PO4, show=c("min", "max", "mean"))
```

```{r}
WaterData %>%
  get_summary_stats(DIN.SRP, show= c("min", "max", "mean"))
```

```{r}
TreatmentWaterData %>%
 
  get_summary_stats(DIN.SRP, show= c("min", "max", "mean"))

```

```{r}
HealthyWaterData %>%
  group_by(Treatment) %>%
  get_summary_stats(DIN.SRP, show= c("min", "max", "mean"))
```

```{r}
Exp.TreatmentWaterData %>%
  group_by(TreatmentDays) %>%
  get_summary_stats(DIN.SRP, show= c("min", "max", "mean"))
```

```{r}
WaterData %>%
  group_by(Group,Timepoint) %>%
  get_summary_stats(NH4, type = "mean_sd")
```

#Visualizing the data

```{r}
NH4.bxp <- ggboxplot(WaterData, x="Timepoint", y="NH4", fill="Group")+
  geom_vline(xintercept= as.factor("TP6"), linetype=4, color="black")+
#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("TP6"), label="End Treatment", y=1.2), colour="black",angle=90, vjust=1.5) +
  geom_vline(xintercept= as.factor("TP1"), linetype=4, color="black")+
#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("TP1"), label="Start Treatment", y=1.2), colour="black",angle=90, vjust=1.5)
NH4.bxp
ggsave("Figures/NH4.bxp.JPG", plot = NH4.bxp, height = 5, width = 9, units = 'in', dpi = 600)

```

```{r}
PO4.bxp <- ggboxplot(WaterData, x="Timepoint", y="PO4", fill="Group") +
  geom_vline(xintercept= as.factor("TP6"), linetype=4, color="black")+
#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("TP6"), label="End Treatment", y=.11), colour="black",angle=90, vjust=1.5) +
  geom_vline(xintercept= as.factor("TP1"), linetype=4, color="black")+
#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("TP1"), label="Start Treatment", y=.11), colour="black",angle=90, vjust=1.5)
PO4.bxp
ggsave("Figures/PO4.bxp.JPG", plot = PO4.bxp, height = 5, width = 9, units = 'in', dpi = 600)

```

```{r}
FilteredWaterData <- filter(WaterData, !Group == "Healthy", !Treatment == "No")
```

```{r}
FilteredWaterData$Group <- factor(FilteredWaterData$Group,levels=c("Experimental", "Control"))
```

```{r}
FilteredWaterData$TreatmentDays <- factor(FilteredWaterData$TreatmentDays, levels = c("Pre-Amendment", "1 hr", "2 Day", "3 Day", "5 Day", "7 Day"))
levels(FilteredWaterData$TreatmentDays)
```

```{r}
DIN.bxp.Filtered <- ggboxplot(FilteredWaterData, x="TreatmentDays", y="DIN", color="Group") +
    scale_color_manual(values=c('#661100',"#999933"), labels=c("Nutrient Amended", "Control")) +
 
  #scale_x_discrete(limits=c("2021-09-24","2021-09-27", "2021-09-30","2021-10-05", "2021-10-10","2021-10-15","2021-10-22"), labels=c("Sept 24", "Sept 27","Sept 30","Oct 5", "Oct 10","Oct 15","Oct 22")) +
   labs(y="DIN Levels (µM)", x= "Time Since Nutrient Deployment")  +

  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"),
) 

DIN.bxp.Filtered
ggsave("Figures/DIN.bxp.Filtered.JPG", plot = DIN.bxp.Filtered, height = 5, width = 9, units = 'in', dpi = 600)

```

#Identifying Outliers \*\*\* We see that there are outliers

```{r}
WaterData%>%
  group_by(Timepoint, Group) %>%
identify_outliers(NH4)
```

#Homogenity of variance

```{r}
WaterData$Treatment <- as.factor(WaterData$Treatment)
WaterData$Timepoint <- as.factor(WaterData$Timepoint)
WaterData$Group <- as.factor(WaterData$Group)
```

*There was homogeneity of variances, as assessed by Levene's test (p \>
0.05).*

```{r}
NutrientWater1 %>%
  group_by(Timepoint) %>%
  levene_test(NH4 ~ Group)
```

#homogenity of covariance Assumption \*\*\* There was homogeneity of
covariances, as assessed by Box's test of equality of covariance
matrices (p \> 0.001)

```{r}
box_m(WaterData[, "NH4", drop = FALSE], WaterData$Group)
```

```{r}

box_m(WaterData[, "NH4_log", drop = FALSE], WaterData$Group)
```

#Running Stats ***We will run a two-way mixed ANOVA*** We see
significant differences between timepoints for all three nutrient
measurement

#filtering out the dates that did not have nutrients

```{r}
NutrientWater <- WaterData %>%
  filter(Timepoint != "T0", Timepoint != "T8", Timepoint != "T9")
NutrientWater
NutrientWater1 <- NutrientWater %>% filter(Group !="Healthy")
NutrientWater1
```

```{r}
NutrientWater1$NH4_log<- log(NutrientWater1$NH4)

```

```{r}
NutrientWater1 %>%
group_by(Group,Timepoint) %>%
  shapiro_test(NH4_log)
```

# Differences with only nutrient timepoints can only look at differences between control and experimental, I believe the healthy and control have too many sililarities

*We can see there is signiicant effect of timepoint of nitrate and
nitrite concentrations* F(6, 48) = 115.092, p \< 0.0001

```{r}
NH4.Mixed.ANOVA <- anova_test(data = NutrientWater1, dv = NH4, wid = CattleTag, within = Timepoint, between = Group)

get_anova_table(NH4.Mixed.ANOVA)

```

#Post-Hoc Tests Pairwise Comparisson between Timepoint of nitrate and
nitrite

```{r}
NH4.pwc <- NutrientWater1 %>%
 
  pairwise_t_test(NH4 ~ Timepoint, p.adjust.method = "bonferroni") 
  
NH4.pwc


```

#Differences with all timepoints

Using the Waterdata dataframe Need to log transform nutrients first

```{r}
WaterData$NH4_log<- log(WaterData$NH4)
WaterData
```

\*F=47.7422

```{r}
NH42.ANOVA <- anova_test(data=WaterData, NH4 ~ Timepoint*Group, wid=CattleTag)
                      get_anova_table(NH42.ANOVA)
```

```{r}

NH4Log.ANOVA <- anova_test(data=WaterData, NH4_log ~ Timepoint*Group, wid=CattleTag)
                      get_anova_table(NH4Log.ANOVA)
```

------------------------------------------------------------------------

#Pairwise interactions of changes in

```{r}
set.seed(999)
NH4.adonis <-  pairwise.adonis(WaterData[c(10)], factors=WaterData$Date, sim.method='euclidian',p.adjust.m='bonferroni', perm = 9999)
NH4.adonis
```

#Phosphate

```{r}

descdist(WaterData$PO4)
```

```{r}
NutrientWater1$PO4_log<- log(NutrientWater1$PO4)
NutrientWater1
```

```{r}
NutrientWater1 %>%
group_by(Group,Timepoint) %>%
  shapiro_test(PO4)
```

#Mixed ANOVA for phosphate levels *Significant effect of time on
phosphate levels* F(6,48) = 13.292, p \<0.001

```{r}
PO4.Mixed.ANOVA <- anova_test(data = NutrientWater1, dv = PO4, wid = CattleTag, within = Timepoint, between = Group)

get_anova_table(PO4.Mixed.ANOVA)
```

```{r}
PO4.pwc <- NutrientWater1 %>%
 
  pairwise_t_test(PO4 ~ Timepoint, p.adjust.method = "bonferroni") 
PO4.pwc

```

#Ammonium

```{r}

descdist(WaterData$NO23)
```

```{r}
NutrientWater1$NO23_log<- log(NutrientWater1$NO23)
NutrientWater1
```

#Mixed ANOVA for ammonium

*Significant diffence between timepoints on ammonium levels F(6,48) =
5.035, p \< 0.005*

```{r}
NO23.Mixed.ANOVA <- anova_test(data = NutrientWater1, dv = NO23, wid = CattleTag, within = Timepoint, between = Group)
NO23.Mixed.ANOVA
get_anova_table(NO23.Mixed.ANOVA)
```

#Pairwise comparisons of ammonium differences

```{r}
NO23.pwc <- NutrientWater1 %>%
 
  pairwise_t_test(NO23 ~ Timepoint, p.adjust.method = "bonferroni") 
NO23.pwc

```



```{r}
NO23.bxp <- ggboxplot(WaterData, x="Timepoint", y="NO23", fill="Group") +
  geom_vline(xintercept= as.factor("TP6"), linetype=4, color="black")+
#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("TP6"), label="End Treatment", y=1.5), colour="black",angle=90, vjust=1.5) +
  geom_vline(xintercept= as.factor("TP1"), linetype=4, color="black")+
#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("TP1"), label="Start Treatment", y=1.5), colour="black",angle=90, vjust=1.5)
NO23.bxp
ggsave("Figures/NO23.bxp.JPG", plot = NO23.bxp, height = 5, width = 9, units = 'in', dpi = 600)

```

#DIN.SRP

```{r}

descdist(WaterData$DIN.SRP)
```

```{r}
WaterData$DIN.SRP_log<- log(WaterData$DIN.SRP)
WaterData
```



#Check DIN.SRP Values for Normalicy \*\*\* All data is normal except for
two of the timepoints between groups

```{r}
GroupWaterData %>%
  group_by(Group,Timepoint) %>%
  shapiro_test(DIN.SRP)
```

------------------------------------------------------------------------

Check Summary Statistics

```{r}
GroupWaterData %>%
  group_by(Group,Timepoint) %>%
  get_summary_stats(DIN.SRP, type = "mean_sd")
```

bxp \<- ggboxplot(WaterData, x="Timepoint", y="NH4", fill="Group") bxp
ggsave("NH4.bxp.JPG", plot = bxp, height = 9, width = 9, units = 'in',
dpi = 600)

```{r}
DIN.SRP.Bxp <- ggboxplot(GroupWaterData, x="Timepoint", y="DIN.SRP", fill="Group") +
  geom_vline(xintercept= as.factor("TP6"), linetype=4, color="black")+
#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("TP6"), label="End Treatment", y=20), colour="black",angle=90, vjust=1.5) +
  geom_vline(xintercept= as.factor("TP1"), linetype=4, color="black")+
#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("TP1"), label="Start Treatment", y=20), colour="black",angle=90, vjust=1.5)
DIN.SRP.Bxp
ggsave("Figures/DIN.SRP.bxp.JPG", plot = DIN.SRP.Bxp, height = 5, width = 9, units = 'in', dpi = 600)
```

------------------------------------------------------------------------

Identifying Outliers \*\*\* We can see there are a few extreme outliers,
also shown on the graph above, therefore we can assume the data is
non-normal and proceed with PERMANOVA tests

```{r}
GroupWaterData%>%
  group_by(Timepoint, Group) %>%
  identify_outliers(DIN.SRP)
```

#Running Mixed ANOVA on DIN.SRP \*\*\* Found Significant differences
between timepoints F(6,48) = 8.720, p \< 0.0001

```{r}
DIN.SRP.Mixed.ANOVA <- anova_test(data = NutrientWater1, dv = DIN.SRP, wid = CattleTag, within = Timepoint, between = Group)

get_anova_table(DIN.SRP.Mixed.ANOVA)


```
```{r}

DIN.SRP.pwc <- NutrientWater1 %>%
 
  pairwise_t_test(DIN.SRP ~ Timepoint, p.adjust.method = "bonferroni") 
DIN.SRP.pwc


```

##ANALIZING AMBIENT WATER DATA THROUGH TIME

#Check normalcy \*\*\* We can see the data is normally distribulted
through shapiro wilk tst

```{r}
HealthyWaterData %>%
  group_by(Timepoint) %>%
  shapiro_test(DIN.SRP)
```

#Visualizing the data

```{r}
Ambientbxp <-  ggboxplot(HealthyWaterData, x="Timepoint", y="DIN.SRP")
Ambientbxp
```

#Running statistical tests

Because the data was normally distributed we will run an ANOVA_test

```{r}
Ambient.ANOVA <- anova_test(data=HealthyWaterData, DIN.SRP ~ Timepoint, wid=CattleTag)
get_anova_table(Ambient.ANOVA)
```

```{r}
HealthyWaterData %>%
pairwise_t_test(DIN.SRP~Date, paired=FALSE, p.adjust.method="bonferroni") 

```

#LOOKING AT TREATMENT DIFFERENCES

------------------------------------------------------------------------

Now we want to see if there were any differences from when the nutrients
were present on the *Experimental* colonies and when they were not We
will use the *TreatmentWaterData* for this, where the only group is
*Experimental* and we will see if there is a significant interaction
between DIN.SRP levels between *Treatment* which is yes or no

#Check noramility assumptions \*\*\* The data is not normally
distributed

```{r}
TreatmentWaterData %>%
  group_by(Treatment) %>%
  shapiro_test(DIN.SRP)
```

#Visualize the data

```{r}
Treatment.DIN.Bxp <-  ggboxplot(TreatmentWaterData, x="Treatment", y="DIN.SRP")
Treatment.DIN.Bxp
```

#Running a statistical non-parametric equivalent to a one-way ANOVA ***-
Insignificant effect of treatment on DIN:SRP ratios (Kruskal-Wallis: H =
3.018, p \< 0.082)***

```{r}
Treatment.Kruskal <- TreatmentWaterData %>% kruskal_test(DIN.SRP ~ Treatment)
Treatment.Kruskal
TreatmentWaterData %>% kruskal_effsize(DIN.SRP ~ Treatment) 
  
```

#Looking at changes in nutrient amount over time

------------------------------------------------------------------------

We filtered out the timepoints that were not during nutrient ammendments

```{r}
NutrientWaterData <- filter(TreatmentWaterData, Treatment=="Yes")
NutrientWaterData
```

```{r}
NutrientWaterData %>%
  group_by(TreatmentDays) %>%
  shapiro_test(DIN.SRP)
```

#Testing to see if there are any significant differences in DIN:SRP
ratios between nubmer of treatment days

```{r}
TreatmentDays.Kruskal <- NutrientWaterData %>% kruskal_test(DIN.SRP ~ TreatmentDays)
TreatmentDays.Kruskal

```

```{r}
NH4TreatmentDays.Kruskal <- NutrientWaterData %>% kruskal_test(NH4 ~ TreatmentDays)
NH4TreatmentDays.Kruskal

```

# Visualizing the data

------------------------------------------------------------------------

We are visualizing if there are any effects of the number of days on
each ofthe nutrient levels over the nubmer of days since treatment. From
here we can see there were no differences in the amoutn of days since
nutrient levels except in the very beginning.

```{r}
Nutrient.bxp <-  ggboxplot(NutrientWaterData, x="TreatmentDays", y="DIN.SRP")
Nutrient.bxp
```

```{r}
Nutrient.bxp1 <-  ggboxplot(NutrientWaterData, x="TreatmentDays", y="NO23")
Nutrient.bxp1
```

```{r}
Nutrient.bxp2 <-  ggboxplot(NutrientWaterData, x="TreatmentDays", y="NH4")
Nutrient.bxp2
```

```{r}
WaterData
HealthyWaterData
mutate(WaterData,DIN = (NO23 + NH4))
Nutrients <- HealthyWaterData %>%
  group_by(Date) %>%
  summarize(PO4=mean(PO4),PO4.SD=sd(PO4),DIN.SD= sd(DIN), DIN=mean(DIN),NH4=mean(NH4), NO23=mean(NO23), sd=sd(DIN.SRP), DIN.SRP=mean(DIN.SRP))

#summarySE(HealthyWaterData, measurevar= c("PO4"), groupvars=c("Date", "Treatment"))
  
Nutrients

```

```{r}
 summarize(HealthyWaterData, mean=mean(DIN))
```

```{r}
DIN.SRP.LineGraph<-  ggplot(data=Nutrients, aes(x=(Date),y=DIN.SRP, color='character')) +
  geom_line( group='character', color="#117733", size=1.5)+
# had to add this grouping so i could change the color
  geom_errorbar(aes(ymin=DIN.SRP-sd, ymax=DIN.SRP+sd), color= "#117733", size=1.5, width=0.5) +
 # scale_color_manual(values =c("117733")) +
  #cannot get the color to work

    
 geom_vline(xintercept= as.factor("2021-09-24"), linetype=4, color="black")+
  geom_text(aes(x=as.factor("2021-09-24"), label="Start Treatment", y=10), colour="black",angle=90, vjust=2) +
  


  geom_vline(xintercept= as.factor("2021-10-22"), linetype=4, color="black")+
geom_text(aes(x=as.factor("2021-10-22"), label="End Treatment", y=10), colour="black",angle=90, vjust=1.5) +


scale_x_discrete(limits=c("2021-09-22", "2021-09-24","2021-09-27", "2021-09-30","2021-10-05", "2021-10-10","2021-10-15","2021-10-22", "2021-11-03", "2021-12-01"), labels=c("Sept22","Sept 24", "Sept 27","Sept 30","Oct 5", "Oct 10","Oct 15", "Oct 22", "Nov 3", "Dec 1")) +
   labs(y="DIN:SRP Ratio (µM)", x= "Date")  +

  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"),
          legend.position = "none") 
DIN.SRP.LineGraph
#ggsave("Figures/DIN.SRP.LineGraph.JPG", plot = DIN.SRP.LineGraph, height = 5, width = 9, units = 'in', dpi = 600)
```

```{r}
DIN.LineGraph<-  ggplot(data=Nutrients, aes(x=(Date),y=DIN, color='character')) +
  geom_line( group='character', color="Blue", size=1.5)+
# had to add this grouping so i could change the color
  geom_errorbar(aes(ymin=DIN-DIN.SD, ymax=DIN+DIN.SD), color= "Blue", size=1.5, width=0.5) +
 # scale_color_manual(values =c("117733")) +
  #cannot get the color to work

    
geom_vline(xintercept= as.factor("2021-09-24"), linetype=4, color="black")+
 # geom_text(aes(x=as.factor("2021-09-24"), label="Start Treatment", y=10), colour="black",angle=90, vjust=2) +
  


  geom_vline(xintercept= as.factor("2021-10-22"), linetype=4, color="black")+
#geom_text(aes(x=as.factor("2021-10-22"), label="End Treatment", y=10), colour="black",angle=90, vjust=1.5) +


scale_x_discrete(limits=c("2021-09-22", "2021-09-24","2021-09-27", "2021-09-30","2021-10-05", "2021-10-10","2021-10-15","2021-10-22", "2021-11-03", "2021-12-01"), labels=c("Sept22","Sept 24", "Sept 27","Sept 30","Oct 5", "Oct 10","Oct 15", "Oct 22", "Nov 3", "Dec 1")) +
   labs(y="DIN (µM)", x= "Date")  +

  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"),
          legend.position = "none") 
DIN.LineGraph
ggsave("Figures/DIN.LineGraph.JPG", plot = DIN.LineGraph, height = 5, width = 9, units = 'in', dpi = 600)
```

```{r}
#Nutrient.Data$Date <- factor(NutrientData$Date, labels=c(c("Sept 22","Sept24","Sept27","Sept 30","Oct5", "Oct 10","Oct15","Oct 22", "Nov 3", "Dec 1"))) 
FilteredOrderedData <- filter(OrderedNutrientData, Nutrient %in% c("PO4", "NO23", "NH4"))
OrderedNutrientData.Summary1 <- summarySE(FilteredOrderedData, measurevar="Measurement", groupvars = c("Nutrient", "Date", "Group"), na.rm=TRUE)
OrderedNutrientData.Summary1
```

```{r}


OrderedNutrientData.Summary1$Group <- factor(OrderedNutrientData.Summary1$Group,levels=c("Experimental", "Control", "Healthy"))


```


```{r}
AllNutrient.LineGraph <- ggplot(OrderedNutrientData.Summary1, aes(x=Date,y=Measurement,color=Group, group=Group)) +
  geom_line(size=0.8)+
#geom_bar(stat="identity", position="dodge")+
  geom_point()+
  geom_errorbar(aes(ymin=Measurement-se, ymax=Measurement+se, group=Group), width =0.2) + 
   scale_color_manual(values=c('#661100',"#999933", "#6699CC"), labels=c("Nutrient Amended", "Control", "Healthy")) +
   facet_wrap(Nutrient~., scales="free_y", strip.position="right", ncol=1) +
 # geom_vline(xintercept= as.factor("2021-09-24"), linetype=4, color="black")+
  #geom_text(aes(x=as.factor("2021-09-30"), label="Start Treatment", y=1), colour="black",angle=90, vjust=1.5) +
  


 # geom_vline(xintercept= as.factor("2021-10-22"), linetype=4, color="black")+
#geom_text(aes(x=as.factor("2021-10-22"), label="End Treatment"), colour="black",angle=90, vjust=1.5) +
  



scale_x_discrete(limits=c("2021-09-22", "2021-09-24","2021-09-27", "2021-09-30","2021-10-05", "2021-10-10","2021-10-15","2021-10-22", "2021-11-03", "2021-12-01"), labels=c("Sept22","Sept 24", "Sept 27","Sept 30","Oct 5", "Oct 10","Oct 15", "Oct 22", "Nov 3", "Dec 1")) +
   labs(y="Nutrient Concentration (µM)")  +

  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"), 
        legend.key=element_blank())
AllNutrient.LineGraph
ggsave("Figures/AllNutrients.LineGraph.JPG", plot =AllNutrient.LineGraph, height = 9, width = 9, units = 'in', dpi = 600)
```

#Local Rainfall Data

```{r}
HealthyWaterData %>%
  shapiro_test(Rain)
```

```{r}
RainData1 <- HealthyWaterData %>%
  group_by(Date) %>%
  summarize(meanRain = mean(Rain), meanDIN = mean(DIN.SRP), meanNH4= mean(NH4), meanPO4= mean(PO4), meanNO23=mean(NO23))
RainData1

```

```{r}
shapiro_test(RainData, meanRain)
```

```{r}

cor.test(HealthyWaterData$Rain, HealthyWaterData$DIN.SRP, method="pearson")
```

```{r}
cor.test(RainData$meanRain, RainData$meanDIN, method="pearson")
```

#Trying to visualize the effect of rainfall on nutrient levels ***From
correlation analysis there is no relationship between average rainfall
and DIN.SRP ratios, there also doesn't visually look like there are any
differences***

```{r}
ggplot(RainData, aes(x=Date, y=meanRain, group=1)) +
  geom_line()
```

```{r}
cor.test()
```

```{r}
ggplot(RainData, aes(x=Date, group=1)) +
```

```{r}
geom_bar(aes(y=Rainfall), stat="identity") +

```

```{r}
RainData
```

```{r}
ggplot(RainData, aes(x=Date, group=1)) +
geom_point(color="blue", (aes(y=Type))) +

  geom_point(aes(y=NO23, color="blue")) +
  geom_point(color="green", aes(y=PO4))

```

```{r}
HealthyWaterData
```

```{r}
RainData2 <- filter(RainData, Type != "DIN.SRP", Type != "PO4")
RainData2

```

```{r}
ggplot(RainData2, aes(x=Date, y=Level, group=Type, color=Type)) +
  geom_line()+
  geom_point() +
    scale_x_discrete(breaks=c("2021-09-22", "2021-09-24","2021-09-27", "2021-09-30","2021-10-05", "2021-10-10","2021-10-15","2021-10-22", "2021-11-03", "2021-12-01"), labels=c("Sept22","Sept 24", "Sept 27","Sept 30","Oct 5", "Oct 10","Oct 15", "Oct 22", "Nov 3", "Dec 1")) 
```

#Figures and Stats for NEW nutrient data

```{r}
NewWaterDataDIN$Colony <- as.factor(NewWaterDataDIN$Colony)
NewWaterDataDIN$Date <- as.factor(NewWaterDataDIN$Date)
NewWaterDataDIN$Group <- as.factor(NewWaterDataDIN$Group)
```

```{r}
#Nutrient.Data.SD <- summarySE(OrderedNutrientData, measurevar= "Measurement", groupvars=c("Nutrient", "Group", "Date", "Treatment"), na.rm=TRUE)
#Nutrient.Data.SD

NewNutrients.SD <- summarySE(NewWaterData, measurevar = "Measurement", groupvars=c("Nutrient", "Group", "Date"), na.rm=TRUE) 
NewNutrients.SD


NewWaterData$MeasurementLog <- log(NewWaterData$Measurement)
NewNutrients.Log <- summarySE(NewWaterData, measurevar = "MeasurementLog", groupvars=c("Nutrient", "Group", "Date"), na.rm=TRUE) 
NewNutrients.Log
```

```{r}
NewNutrients.SD$Group <- factor(NewNutrients.SD$Group,levels=c("Nutrient Ammended", "Control"))
NewNutrients.Log$Group <- factor(NewNutrients.Log$Group,levels=c("Nutrient Ammended", "Control"))
```

```{r}
#Nutrient.Data$Nutrient <- factor(Nutrient.Data$Nutrient, levels=c("NH3","NO23","PO4"), labels=c("NH4", "NO23", "PO4"))



NH4bxp <- ggboxplot(NewWaterDataDIN, x="Date", y="NH4_Log", fill="Group")
NH4bxp
```

```{r}
NewNutrients.LineGraph <- ggplot(NewNutrients.SD, aes(x=Date,y=Measurement,color=Group, group=Group)) +
  geom_line(size=0.8)+
#geom_bar(stat="identity", position="dodge")+
  geom_point()+
  geom_errorbar(aes(ymin=Measurement - se, ymax= Measurement + se, group = Group), width =0.2) +
   scale_color_manual(values=c('#661100', "#999933" ), labels=c("Nutrient Amended", "Control")) +
   facet_wrap(Nutrient~., scales="free_y", strip.position="right", ncol=1) +
 # geom_vline(xintercept= as.factor("2021-09-24"), linetype=4, color="black")+
  #geom_text(aes(x=as.factor("2021-09-30"), label="Start Treatment", y=1), colour="black",angle=90, vjust=1.5) +
  


 # geom_vline(xintercept= as.factor("2021-10-22"), linetype=4, color="black")+
#geom_text(aes(x=as.factor("2021-10-22"), label="End Treatment"), colour="black",angle=90, vjust=1.5) +
  



scale_x_discrete(limits=c("2023-03-05", "2023-03-06","2023-03-08", "2023-03-10"), labels=c("1 hr","1 day", "2 days","4 days")) +
   labs(y="Trial 2 Nutrient Concentrations (µM)", x ="Timepoint")  +

  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"), 
        legend.key=element_blank())
NewNutrients.LineGraph
ggsave("Figures/NewAllNutrients.LineGraph.JPG", plot =NewNutrients.LineGraph, height = 6, width = 11, units = 'in', dpi = 600)
```
```{r}
NewNutrients.Log_LineGraph <- ggplot(NewNutrients.Log, aes(x=Date,y=MeasurementLog,color=Group, group=Group)) +
  geom_line(size=0.8)+
#geom_bar(stat="identity", position="dodge")+
  geom_point()+
  geom_errorbar(aes(ymin=MeasurementLog - se, ymax= MeasurementLog + se, group = Group), width =0.2) +
   scale_color_manual(values=c('#661100', "#999933" ), labels=c("Nutrient Amended", "Control")) +
   facet_wrap(Nutrient~., scales="free_y", strip.position="right", ncol=1) +
 # geom_vline(xintercept= as.factor("2021-09-24"), linetype=4, color="black")+
  #geom_text(aes(x=as.factor("2021-09-30"), label="Start Treatment", y=1), colour="black",angle=90, vjust=1.5) +
  


 # geom_vline(xintercept= as.factor("2021-10-22"), linetype=4, color="black")+
#geom_text(aes(x=as.factor("2021-10-22"), label="End Treatment"), colour="black",angle=90, vjust=1.5) +
  



scale_x_discrete(limits=c("2023-03-05", "2023-03-06","2023-03-08", "2023-03-10"), labels=c("1 hr","1 day", "2 days","4 days")) +
   labs(y="Trial 2 Nutrient Concentrations (µM)", x ="Timepoint")  +

  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"), 
        legend.key=element_blank())
NewNutrients.Log_LineGraph
ggsave("Figures/NewAllNutrients.Log.LineGraph.JPG", plot =NewNutrients.Log_LineGraph, height = 6, width = 11, units = 'in', dpi = 600)
```

#NEW Nitrate & Nitrite

*No signficant difference betwee timepoint or group (two-way ANOVA:
F32,3=2.699, p= 0.062*

```{r}
NewWaterDataDIN %>%
group_by(Group,Date) %>%
  shapiro_test(NH4)
NewWaterDataDIN %>%
group_by(Group,Date) %>%
  shapiro_test(NH4_Log)
```

```{r}
box_m(NewWaterDataDIN[, "NH4", drop = FALSE], NewWaterDataDIN$Group)
```

```{r}

NewWaterDataDIN %>%
  group_by(Date)%>%
  levene_test(NH4 ~ Group)
descdist(NewWaterDataDIN$NH4)
descdist(NewWaterDataDIN$NH4_Log)
```

```{r}
set.seed(999)
PERMA.NewNH4<- adonis2(NewWaterDataDIN$NH4~Group*Date, data=NewWaterDataDIN)
PERMA.NewNH4
```

#Two Way Mixed ANOVA on new Nitrate and Nitrate concentraitons
Significant two-way interactions between group and time on anxiety score
F(3,24) = 4.044, P = 0.018

```{r}

NewWaterDataDIN$NH4_Log<- log(NewWaterDataDIN$NH4)

New_NH4.ANOVA <- anova_test(data = NewWaterDataDIN, dv = NH4, wid = Colony, within = Date, between = Group)

get_anova_table(New_NH4.ANOVA)
New_NH4_Log.ANOVA <- anova_test(data = NewWaterDataDIN, dv = NH4_Log, wid = Colony, within = Date, between = Group)

get_anova_table(New_NH4_Log.ANOVA)
```

#Effect of NH4 group at each timepoint Significant main effect of group
was siginificant the one hour timepoint (P = 0.021)

```{r}
New.NH4.OneWay <- NewWaterDataDIN%>%
  group_by(Date) %>%
  anova_test(dv =NH4_Log, wid = Colony, between = Group) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
New.NH4.OneWay
```

#NH4 Pairwise comparisons between group levels Significant difference of
control and nutrient groups at the one hour tiepoint (p = 0.021)

```{r}
NewNH4.pwc <- NewWaterDataDIN %>%
  group_by(Date) %>%
  pairwise_t_test(NH4_Log ~ Group, p.adjust.method = "bonferroni")
NewNH4.pwc
```

#NH4 Effect of time
```{r}
New.NH4.OneWayTime <- NewWaterDataDIN%>%
  group_by(Group) %>%
  anova_test(dv =NH4_Log, wid = Colony, between = Date) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
New.NH4.OneWayTime
```


```{r}
NewNH4.Time.pwc <- NewWaterDataDIN %>%
  group_by(Group) %>%
  pairwise_t_test(NH4_Log ~ Date, paired=TRUE, p.adjust.method = "bonferroni") %>%
  select(-df, -statistic, -p) # Remove details
NewNH4.Time.pwc
```

```{r}
NewWaterDataDIN %>%
  pairwise_t_test(NH4 ~ Timepoint, paired=TRUE, p.adjust.method="bonferroni") 
```

```{r}
NewDIN.bxp <- ggboxplot(NewWaterDataDIN, x="Timepoint", y="DIN", color="Group") +
    scale_color_manual(values=c('#661100',"#999933"), labels=c("Nutrient Amended", "Control")) +
 
  #scale_x_discrete(limits=c("2021-09-24","2021-09-27", "2021-09-30","2021-10-05", "2021-10-10","2021-10-15","2021-10-22"), labels=c("Sept 24", "Sept 27","Sept 30","Oct 5", "Oct 10","Oct 15","Oct 22")) +
   labs(y="DIN Levels (µM)", x= "Treatment Dates")  +

  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"),
) 

NewDIN.bxp
#ggsave("Figures/DIN.bxp.JPG", plot = DIN.bxp, height = 5, width = 9, units = 'in', dpi = 600)

```

From the graph above there looks to be outliers that could be driving
the graph, so we are going to identify, remove and determine the
difference without

```{r}

NewWaterDataDIN%>%
  group_by(Date, Group) %>%
identify_outliers(DIN)
```

```{r}
FilteredOutliersNew <- 
  filter(NewWaterDataDIN, !DIN > 70, !DIN == 0.915)
FilteredOutliersNew
```

```{r}

#FilteredOutliersNew$Group <- factor(FilteredOutliersNew$Group,levels=c("Nutrient Amended", "Control"))
FilteredOutliersNew$Timepoint <- factor(FilteredOutliersNew$Timepoint, levels = c("Pre-Amendment", "1hr", "1 Day", "3 Day", "5 Day"))
```

```{r}
FilteredNewDIN.bxp <- ggboxplot(FilteredOutliersNew, x="Timepoint", y="DIN", color="Group") +
    scale_color_manual(values=c("#999933",'#661100'), labels=c("Control", "Nutrient Amendment")) +
 
  #scale_x_discrete(limits=c("2021-09-24","2021-09-27", "2021-09-30","2021-10-05", "2021-10-10","2021-10-15","2021-10-22"), labels=c("Sept 24", "Sept 27","Sept 30","Oct 5", "Oct 10","Oct 15","Oct 22")) +
   labs(y="DIN Levels (µM)", x= "Time Since Nutrient Deployment")  +

  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"),
) 

FilteredNewDIN.bxp
ggsave("Figures/FilteredNewDIN.bxp.JPG", plot = FilteredNewDIN.bxp, height = 5, width = 9, units = 'in', dpi = 600)
```

```{r}
set.seed(999)
PERMA.NewNH4<- adonis2(NewWaterDataDIN$NH4~Group*Date, data=NewWaterDataDIN)
PERMA.NewNH4
```

#Ammonium

```{r}
NewWaterDataDIN$NO23_Log <- log(NewWaterDataDIN$NO23)

NewWaterDataDIN %>%
group_by(Group,Date) %>%
  shapiro_test(NO23)

NewWaterDataDIN %>%
group_by(Group,Date) %>%
  shapiro_test(NO23_Log)

```

```{r}


descdist(NewWaterDataDIN$NO23)
```

#Running Mixed ANOVA on ammonium

*Significant difference between date, group, and time (mixed ANOVA:
F3,24 F=4.672, p=0.010)*

```{r}
New_NO23_Log.ANOVA <- anova_test(data = NewWaterDataDIN, dv = NO23_Log, wid = Colony, within = Date, between = Group)

get_anova_table(New_NO23_Log.ANOVA)
```

Post-hoc test NO23

```{r}
New.NO23.OneWay <- NewWaterDataDIN %>%
  group_by(Date) %>%
  anova_test(dv =NO23_Log, wid = Colony, between = Group) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
New.NH4.OneWay
```

#NO23 Pairwise comparisons between group levels Significant difference
of control and nutrient groups at the one hour tiepoint (p = 0.021)

```{r}
NewNO23.pwc <- NewWaterDataDIN %>%
  group_by(Date) %>%
  pairwise_t_test(NO23_Log ~ Group, p.adjust.method = "bonferroni")
NewNO23.pwc
```

#NO23 Effect of time

```{r}
NewNO23.Time.pwc <- NewWaterDataDIN %>%
  group_by(Group) %>%
  pairwise_t_test(NO23_Log ~ Date, paired=TRUE, p.adjust.method = "bonferroni") %>%
  select(-df, -statistic, -p) # Remove details
NewNO23.Time.pwc
```

#Phosphate *Significant difference between date, and interaction of date
& group*

```{r}
descdist(NewWaterDataDIN$PO4)
```

```{r}

NewWaterDataDIN$PO4_Log<- log(NewWaterDataDIN$PO4)
NewWaterDataDIN %>%
group_by(Group,Date) %>%
  shapiro_test(PO4)

NewWaterDataDIN %>%
group_by(Group,Date) %>%
  shapiro_test(PO4_Log)

```

```{r}
New_PO4_Log.ANOVA <- anova_test(data = NewWaterDataDIN, dv = PO4_Log, wid = Colony, within = Date, between = Group)

get_anova_table(New_PO4_Log.ANOVA)
```

Post-hoc test NO23

#PO4 Effect of time

```{r}
NewPO4.Time.pwc <- NewWaterDataDIN %>%

  pairwise_t_test(PO4_Log ~ Date, paired=TRUE, p.adjust.method = "bonferroni") 
 
NewPO4.Time.pwc
```

#DIN.SRP *Significant difference between group only*

```{r}

descdist(NewWaterDataDIN$DIN.SRP)
```

```{r}
NewWaterDataDIN %>%
group_by(Group,Date) %>%
  shapiro_test(DIN.SRP)
```

#Mixed ANOVA of DIN.SRP Significant effect of group on DIN.SRP levels

```{r}
New_DIN.SRP.ANOVA <- anova_test(data = NewWaterDataDIN, dv = DIN.SRP, wid = Colony, within = Date, between = Group)

get_anova_table(New_DIN.SRP.ANOVA)
```

#DIN.SRP Effect of time

```{r}
NewDIN.SRP.Time.pwc <- NewWaterDataDIN %>%

  pairwise_t_test(DIN.SRP ~ Group,  p.adjust.method = "bonferroni")
 
NewDIN.SRP.Time.pwc
```
```{r}
ggboxplot(NewWaterDataDIN, x="Group", y="DIN.SRP", fill="Group")

```

Subsetting the data by date to see if there are significant differences
between the groups

```{r}
NewWaterDataDIN %>%
group_by(Group,Date) %>%
  shapiro_test(DIN)
```

```{r}
Filter1 <- filter(FilteredOutliersNew, Date == "2023-03-05")
hr1 <- filter(NewWaterDataDIN, Date == "2023-03-05")
Day1 <- filter(NewWaterDataDIN, Date == "2023-03-06") 
Day3 <- filter(NewWaterDataDIN, Date == "2023-03-08")
Day5 <- filter(NewWaterDataDIN, Date == "2023-03-10")
```

```{r}
NewWaterDataDIN %>% anova_test(DIN~Group*Date)
```

```{r}
NewWaterDataDIN %>% group_by(Date) %>%
  anova_test(DIN~Group)
```

```{r}
FilteredOutliersNew %>% group_by(Date) %>%
  emmeans_test(DIN~Group, p.adjust.method = "bonferroni")
```

```{r}
FilteredOutliersNew %>% group_by(Date) %>%
  emmeans_test(NO23~Group, p.adjust.method = "bonferroni")
```

```{r}
FilteredOutliersNew %>% group_by(Date) %>%
  emmeans_test(NH3~Group, p.adjust.method = "bonferroni")
```

```{r}
FilteredOutliersNew %>% group_by(Date) %>%
  emmeans_test(PO4~Group, p.adjust.method = "bonferroni")
```

```{r}
Filterhr1 %>% 

  wilcox_test(DIN~Group) %>%
  add_significance()
```

```{r}
hr1 %>%
    wilcox_test(DIN~Group) %>%
  add_significance()
```

```{r}
Day1 %>%

  wilcox_test(DIN~Group) %>%
  add_significance()
```

stat.test \<- genderweight %\>% wilcox_test(weight \~ group) %\>%
add_significance() stat.test

```{r}
Day3 %>%
  wilcox_test(DIN~Group) %>%
  add_significance()
```

```{r}
Day5 %>% 
  wilcox_test(DIN~Group) %>%
  add_significance()
```
