---
title: "Thesis Data"
author: "Ashley"
date: "2022-08-29"
output: html_document
---

# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 

```{r, packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}

if (!require("pacman")) install.packages("pacman") 
pacman::p_load(flextable, officer, stringr, ggplot2, stringr, gridExtra, ggpubr, Rmisc, FSA, rcompanion, RColorBrewer, dplyr, vegan, nparcomp, RVAideMemoire, MANOVA.RM, pairwiseAdonis, PMCMR, PMCMRplus, patchwork, plyr, gdata,paletteer, RColorBrewer, devtools, ochRe, rstatix, lmer4, lmerTest, glmmTMB,car,broom,emmeans, lubridate, fitdistrplus, microshades)
pacman::p_load_gh("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
```

##Uploading all thesis data from Excel

```{r, load, include = "true"}
ThesisData <- read.xls("AllNutrientData.xlsx", na.string = "NA", head = T)
ActiveProportionData <- read.xls("DiseaseStatusProportions.xlsx")
ExpPhaseModelData <- read.xls("ExpPhaseProportionData.xlsx")
StatusProportionData <- read.xls("ActiveProportions.xlsx")
LesionExpPhaseData <- read.xls("LesionExpPhase.xlsx")
ModelProportionRate <- read.xls("ModelProportionRate.xlsx")

```

# Organizing the data
***
First need to make *CattleTag* and *DiseaseStatus* as factor variables
Also converting date into a more usuable format
-> change format in excel first for lubridate to convert character to date 


```{r}
ThesisData$CattleTag <- as.factor(ThesisData$CattleTag)
ThesisData$DiseaseStatus <- as.factor(ThesisData$DiseaseStatus)
ThesisData$Treatment <- as.factor(ThesisData$Treatment)
ThesisData$Date <- ymd(ThesisData$Date) 
ThesisData$Date <- as.factor(ThesisData$Date)
ModelProportionRate$Date <- ymd(ModelProportionRate$Date)
ModelRateData$CattleTagTimepoint <- as.factor(ModelRateData$CattleTag)

```


#Filtering Data into new data frames
***
We are filtering out the data into new databases to work with only the timepoints that were measured
*StatusData* will contain rates, lesions, proportions, temp, light, and rain 

```{r}
ModelRateData <- filter(ThesisData, TimePoint %in% c("TP2", "TP4", "TP6", "TP7", "TP8"))
LesionRateData <- filter(ThesisData, TimePoint != "TP0")
NutrientData <- filter(ThesisData, TimePoint %in% c("TP0", "TP2", "TP6", "TP8"))
StatusData <- filter(ThesisData, TimePoint %in% c("TP0", "TP2", "TP4", "TP6", "TP7", "TP8"))
ProportionRateData <- filter(ModelProportionRate, TimePoint != "TP0")

```

#Analyzing if there is a significant difference between intial colony sizes between control and experimental

```{r}
ModelAreaTP1 <- filter(ThesisData, TimePoint == "TP0")
```

**Checking Normalicy**
```{r}
ModelAreaTP1 %>% 
  shapiro_test(Area)
```

Can see here that the data is normal

```{r}
descdist(ModelAreaTP1$Area)
```
T test to check if there is a significant difference between the initial colony sizes
*No significant difference between initial colony sizes of experimental and control (t (23.501) = 0.353, p = 0.727)*

```{r}
t.test(Area~Treatment, data=ModelAreaTP1)

```


```{r}
summarise(ModelAreaTP1, mean=mean(Area))
ModelAreaTP1 %>%
  group_by(Treatment) %>%
  summarise(sd=sd(Area), mean=mean(Area))
  
```

 
# Checking Normalicy Assumption for Proportion of Change in healthy tissue area 
***
We are going to check the normalicy of the *AreaProportion*  variables using the *ModelRateData* frame
Here we can see that the Proportion data is normal from the Shapiro Wilks test

*AreaProportion* = (T2-T1)/T2 -> (mmTx^2-mm(T(x-1))^2)/(mm(T(x-1))^2 )

Doing the same for log transformed data
***

```{r}
ModelRateData %>%
group_by(Treatment,TimePoint) %>%
  shapiro_test(AreaProportion)
```

```{r}
descdist(ModelRateData$AreaProportion)
```


```{r}
ModelRateData$AreaProportion_log<- log(ModelRateData$AreaProportion + 1)
ModelRateData
```
**Transformation was attempted to improve normality metrics, but were unsuccessful** 

```{r}
ModelRateData %>%
group_by(Treatment,TimePoint) %>%
  shapiro_test(AreaProportion_log)
```
```{r}
descdist(ModelRateData$AreaProportion_log)
```


***
Visualizing AreaProportion Data
```{r}
ModelRatebxp <- ggboxplot(ModelRateData, x="Date", y="AreaProportion", fill="Treatment")
ModelRatebxp
```
***
Identifying Outliers
```{r}
ModelRateData%>%
  group_by(TimePoint, Treatment) %>%
identify_outliers(AreaProportion)
```

```{r}
ggqqplot(ModelRateData, "AreaProportion") 
```


```{r}

ModelRateData %>%
  group_by(Date) %>%
  levene_test(AreaProportion ~ Treatment)

```


#TWO-WAY MIXED MEASURES ANOVA on Area Proportion

Running a two-way mixed measures ANOVA to see if there is any relationship between the *AreaProportion* (T2-T1)/T1 of tissue change over different *Dates* and between the *Treatment* groups. From the test we can see there is no significant interaction for any of them or between *Treatment* and *TimePoint*. *get-anova_table* the Greenhouse-Geisser sphericity correction is automatically applied to factors violating the sphericity assumption.
This is looking at the proportion of change in healthy tissue area between each timepoint
***Insignificant between treatment and time (F2.7,64.77 = 1.155, p = 0.331,η_G^2 = 0.036 )***
 
```{r}


Proportion.ANOVA <-  
  anova_test(data = ModelRateData, dv = AreaProportion, wid = CattleTag, within= Date,between = Treatment)

get_anova_table(Proportion.ANOVA)



```
 
#Two-Way Repeated Measures Mixed ANOVA Area Proportion over Experiment time
 
 This will use the *ExpPhaseModelData * where we will view the effect of *Group* (experimental or control) and *Treatment* (During treatment, post treatment or total). Proportions were calculated by subtracting the last timepoint during treatment from the timepoint at the start of treatment and dividing by the starting timepoint. (t8-t1)/t1
 Looking at the change in healthy tissue area over the experimental phases
*** 
Checking Normalicy
Non-normal data p < 0.05
Can still run a two-way ANOVA on non-normal data

```{r}
ExpPhaseModelData  %>%
  group_by(Group,Treatment) %>%
  shapiro_test(ExpTissueProportion)

```
```{r}

descdist(ExpPhaseModelData $ExpTissueProportion)
```

 #Log Transformation
 Natural log transformation
 
```{r}

ExpPhaseModelData$ExpTissueProportion_log<- log(ExpPhaseModelData$ExpTissueProportion+1)
ExpPhaseModelData 
```
```{r}
ExpPhaseModelData  %>%
  group_by(Group,Treatment) %>%
  shapiro_test(ExpTissueProportion_log)
```
```{r}
descdist(ExpPhaseModelData $ExpTissueProportion_log)
```

```{r}
ExpPhaseModelData  %>%
 
  levene_test(ExpTissueProportion ~ Treatment*Group)
```
 
 *** 
Runing a two-way anova
***Insignificant between experiment time and treatment group on change in surface area tissue proportion during the experimental phases ((Two-way ANOVA: F72,2 = 0.309, p = 0.735)***
Generating the same results with log transformed data
- We use a two-way anova to evaulate simultaneously the effect of two different grouping variables on an outcome variable 

 
 
```{r}
ExpPhaseAreaProportion.ANOVA <- anova_test(data=ExpPhaseModelData , ExpTissueProportion ~ Treatment*Group)
                   get_anova_table(ExpPhaseAreaProportion.ANOVA)
```

```{r}
ExpPhaseModelData $Group <- factor(ExpPhaseModelData $Group, levels=c("Experimental", "Control"))
                          
                                 
ExpPhaseModelData 
ExpPhaseModelData $Treatment <- factor(ExpPhaseModelData $Treatment, labels=c(c("During Treatment", "Post Treatment", "Total")))


```
```{r}
ExpPhaseModelData$ExpTissueProportionPercent <- ExpPhaseModelData$ExpTissueProportion*100
ExpPhaseModelData
```

```{r}
ModelProportion.Boxplot <- ggplot(data=ExpPhaseModelData , aes(Group,(ExpTissueProportion), color=Group)) +
  geom_violin() +
  geom_jitter(position = position_jitter(width = .2, height=0.07), alpha = 0.6) +#seed = 1, height= 0.055, width = 0.45)) +
  stat_summary(fun.y=median, geom="crossbar", width=.5, alpha=0.5) +


  scale_color_manual(values=c('#661100',"#999933")) +
  scale_x_discrete(labels=c("Nutrient \n Amended", "Control")) +
  facet_wrap(vars(Treatment)) +
   labs(x="Treatment Group", y=expression("Percent of Tissue Change")) +
  scale_y_continuous(labels = scales::percent_format()) +
 #Change the x axis labels
  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"), 
        legend.position = "none")
ModelProportion.Boxplot

ggsave("Figures/Figure7_ModelProportion.Violinplot.JPG", plot = ModelProportion.Boxplot, height = 5, width = 7.7, units = 'in', dpi = 600)
```

#Visualizing Area Proportion change over time for both treatment groups

```{r}
GroupedAreaProportion.Summary <- ModelRateData %>%
  group_by(Date) %>%
  summarise(sd=sd(AreaProportion), AreaProportion=mean(AreaProportion))
GroupedAreaProportion.Summary
AreaProportionTotalSummary <- ModelRateData %>%
  summarise(AreaProportionmean=mean(AreaProportion))
AreaProportionTotalSummary
```

Creating a line graph for *AreaProportion* over *Date* 

```{r}
 AreaProportion.LineGraph <- ggplot(GroupedAreaProportion.Summary,aes(Date,AreaProportion, group=1)) +
  geom_line(color="#AA4499", size=1.5)+
  geom_vline(xintercept= as.factor("2021-10-22"), linetype=4, color="black")+
   geom_hline(yintercept=0, color="black") +
  

#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("2021-10-22"), label="End Treatment", y=-.106), colour="black",angle=90, vjust=2.3)+
  #adds label to intercept line
  # scale_color_manual(values=c('#FC4E07',"#2E9FDF")) +
  geom_errorbar(aes(ymin=AreaProportion-sd, ymax=AreaProportion+sd), color="#AA4499", size=1.5,width=0.5) +

  #adds colors to order of variables


  #scale_color_viridis_c()+
  #scale_fill_viridis_c() +
  #scale_color_manual(values=c('red2','lightcoral')) +
 scale_x_discrete(limits=c( "2021-09-30","2021-10-05", "2021-10-10","2021-10-15","2021-10-22", "2021-11-03", "2021-12-01"), labels=c("Sept 30","Oct 5", "Oct 10","Oct 15", "Oct 22", "Nov 3", "Dec 1")) +
      labs(y="Proportion of Tissue Change" ~(mm^2)) +
 #Change the x axis labels
  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"), 
        legend.position = "none")
      
AreaProportion.LineGraph
ggsave("Figures/AreaProportion.LineGraph.JPG", plot = AreaProportion.LineGraph, height = 5, width = 9, units = 'in', dpi = 600)
```
#TWO-Way ANOVA on Rate of Change in Healthy Tissue Area per day 

Looking overtime so we need to run a two way mixed ANOVA
Testing the effect of treatment and time in the rate of change in tissue area. Using *ProportionRateData* to test *Rate*  (tissue area rate) on *Treatment* and *TimePoint*
***
Check the data and see that the data is non-normal p< 0.05
```{r}
ProportionRateData$Rate <- ProportionRateData$Rate.cm
ProportionRateData %>%
group_by(Treatment,TimePoint) %>%
  shapiro_test(Rate)
```
```{r}

descdist(ModelRateData$Rate)
```
```{r}
ModelRateData %>%
  group_by(TimePoint) %>%
  levene_test(Rate ~ Treatment)
```

 
#Run a GLMM for the Rate data over treatment and date

*** 

No significant difference between *Rate* over *Date* and *Treatment* group

***Insignificant interaction between treatment and timepoint (two way ANOVA: F120,4 = 0.720, p = 0.580)***

#Looking at Proportional rate of change in healthy tissue area per day


This data was found by log(cmTP2/cmTP1)/Days

```{r}
ProportionRateData %>%
group_by(Treatment,TimePoint) %>%
  shapiro_test(ProportionRateLog)
```

```{r}

descdist(ProportionRateData$ProportionRateLog)
```

```{r}

ProportionRateData %>%
  group_by(TimePoint) %>%
  levene_test(ProportionRateLog ~ Treatment)
```

```{r}
LogProportionRates.ANOVA <-  
  anova_test(data = ProportionRateData, dv = ProportionRateLog, wid = CattleTag, within= Date,between =Treatment)

get_anova_table(LogProportionRates.ANOVA)

```

#Visualizing Rate
***
Making a line graph to look at the *AreaRateCm* change over *Time* for each *Treatment* group 
Rate of change in healthy tissue area per day
***
First we need to make summary data

```{r, Error Bar Data}
ModelRateData$AreaRateCm = as.integer(ModelRateData$AreaRateCm)
AreaRate.Summary <- summarySE(ModelRateData, measurevar="AreaRateCm", groupvars = c("Treatment", "Date"), na.rm=TRUE)
AreaRate.Summary


```
***
Reversing the order of the treatment groups to make sure that they are the same order for all graphs

```{r}
AreaRate.Summary$Treatment <- factor(AreaRate.Summary$Treatment, levels=c("Experimental", "Control"))
#AreaRate.Summary$Treatment <- factor(AreaRate.Summary$Treatment, labels=c(c("During Treatment", "Post Treatment", "Total")))
y.axis.title <- bquote(atop(Change~"in"~Healthy~Tissue~Area~phantom(), 
                            Per~Day~(cm^2)))
```

Creating a line graph for *Rate per day* over *Date* between the two *treatments*

```{r}
AreaRate.LineGraph <- ggplot(AreaRate.Summary, aes(Date,AreaRateCm,color=Treatment,group=Treatment)) +
  geom_line(aes(color=Treatment), size=1.5) +
  geom_vline(xintercept= as.factor("2021-10-22"), linetype=4, color="black")+
  geom_hline(yintercept=0, color="black") +
#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("2021-10-22"), label="End Treatment", y=15), colour="black",angle=90, vjust=1.5)+
  #adds label to intercept line
  geom_errorbar(aes(ymin=AreaRateCm-se, ymax=AreaRateCm+se), size = 1.5, width=0.5) +
 scale_color_manual(values=c('#661100',"#999933"),name=("Treatment Group"), labels=c("Nutrient Amended", "Control")) +
  #adds colors to order of variables, and also changes the names 
  #These are colorblind compatible
  #scale_color_discrete(labels=c("Nutrient \n Amended", "Control")) +
  #scale_color_viridis_c()+
  #scale_fill_viridis_c() +
  #scale_color_manual(values=c('red2','lightcoral')) +
  
 scale_x_discrete(limits=c("2021-09-30", "2021-10-10","2021-10-22", "2021-11-03", "2021-12-01"), labels=c("Sept 30", "Oct 10", "Oct 22", "Nov 3", "Dec 1")) +
  labs(y=y.axis.title) +
   #labs(y=expression("Change in Healthy Tissue Area   \n     Per Day" ~(cm^2))) +
    #labs(y="Surface Area Rate of Change \nPer Day (cm^2)") +
 #Change the x axis labels
  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"), 
        legend.position = "none")
       legend.key=element_blank() #Removes the color of the legend

AreaRate.LineGraph
ggsave("Figures/Figure6a_AreaRate.LineGraph.JPG", plot = AreaRate.LineGraph, height = 5, width = 6, units = 'in', dpi = 600)
```

 
#Looking Area Rates differences between During Treatment and Without Treatment 
***
No significant differences
This is looking at the rates for every timepoint grouped with having nutrient or without. This is different than what we looked at earlier with area proportions. 

***-	Insignificant interaction between treatment and fertilizer treatment (Two Way ANOVA: F126,1 = 0.031, p = 0.861)***
*With proportion rates transformed still insiginificant (Two Way ANOVA: F126,1 = 0.034, p = .854)*


```{r}
FertilizerRates.ANOVA <- anova_test(data=ModelRateData, Rate ~ Treatment*Fertilzer)
                    FertilizerRates.ANOVA
```
```{r}
ProportionFertilizerRates.ANOVA <- anova_test(data=ProportionRateData, ProportionRateLog ~ Treatment*Fertilizer)
                      ProportionFertilizerRates.ANOVA
```


# Two WAY repeated measures mixed ANOVA LINEAR LESION RATE DATA
#Check Normalicy 
For these analysis' we will use *LesionRateData* Dataframe to look at if there are any significant differences between *LesionRate* between *Date* and *Treatment* group. This data will be used to validate model data 
Proportional Lesion Rate = Log(T2/T1)/Days
***
Data is not normally distributed, ANOVA should be robust enough 

```{r}
LesionRateData %>%
  group_by(Treatment,Date) %>%
  shapiro_test(LesionRate)

```

```{r}

descdist(LesionRateData$LesionRate)
```

When we try to transform the data it does not make it any more normal 

```{r}
LesionRateData %>%
  group_by(Treatment,Date) %>%
  shapiro_test(ProportionalLesionaRate)

```

```{r}
LesionRateData %>%
  group_by(TimePoint) %>%
  levene_test(ProportionalLesionaRate ~ Treatment)
```



```{r}

descdist(LesionRateData$ProportionalLesionaRate)
```



#Running a mixed TWO-Way ANOVA on Proportional Change of linear lesion rate per day
***Significant interaction between lesion rate and timepoint (Two Way ANOVA: F192,7 = 2.134, p = 0.042)***
```{r}
LesionRate.ANOVA <-  
  anova_test(data = LesionRateData, dv = ProportionalLesionaRate, wid = CattleTag, within= TimePoint,between = Treatment)

get_anova_table(LesionRate.ANOVA)
```

#Post Hoc tests simple pairwise comparison on treatment

```{r}
LesionRateData %>% 
  pairwise_t_test(ProportionalLesionaRate ~ Treatment, p.adjust.method = "bonferroni")
```


**Also want to look at lesion rates grouped by experimental phase (with nutrients and without)**
This will use the *LesionExpPhaseData* looking at the effect of *Treatment* and *ExperimentPhase* on *Measurement*
**(two-way ANOVA: F48,1 = 0.428, p = 0.516)**

```{r}
LesionRateExperimentalPhase.ANOVA <- anova_test(data=LesionExpPhaseData, Measurement ~ Treatment*ExperimentalPhase)
               LesionRateExperimentalPhase.ANOVA
```


#Visualizing Lesion Rate Data
First we need to collect error bar data 

```{r, Error Bar Data}

LesionRate.Summary <- summarySE(LesionRateData, measurevar="LesionRate", groupvars = c("Treatment", "Date"), na.rm=TRUE)
LesionRate.Summary

```

***
Making sure the levels are in the right order for the graphs

```{r}
LesionRate.Summary$Treatment <- factor(LesionRate.Summary$Treatment, levels= c("Experimental", "Control"))

```

Creating a line graph for *Rate per day* over *Date* between the two *treatments*

```{r}
 LesionRate.LineGraph <- ggplot(LesionRate.Summary, aes(Date,LesionRate,color=Treatment,group=Treatment)) +
  geom_line(aes(color=Treatment), size=1.5)+
  geom_vline(xintercept= as.factor("2021-10-22"), linetype=4, color="black")+
#adds intercept lint if not number have to do as.factor or whatever it might be
    geom_hline(yintercept=0, color="black") +
  geom_text(aes(x=as.factor("2021-10-22"), label="End Treatment", y=.3), colour="black",angle=90, vjust=1.5)+
  #adds label to intercept line
  geom_errorbar(aes(ymin=LesionRate-se, ymax=LesionRate+se), linewidth=1.5, width=0.5) +
scale_color_manual(values=c('#661100',"#999933"),name=("Treatment Group"), labels=c("Nutrient Amended", "Control")) +
  #adds colors to order of variables


 scale_x_discrete(limits=c("2021-09-27", "2021-09-30","2021-10-05", "2021-10-10","2021-10-15","2021-10-22", "2021-11-03", "2021-12-01"), labels=c("Sept 27","Sept 30","Oct 5", "Oct 10","Oct 15", "Oct 22", "Nov 3", "Dec 1")) +
      labs(y="Change in Linear Lesion Progression \n Per Day (cm)") +
 #Change the x axis labels
  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"), 
       
        legend.key=element_blank())
   

LesionRate.LineGraph
ggsave("Figures/Figure6b_LesionRate.LineGraph.JPG", plot = LesionRate.LineGraph, height = 5, width = 9, units = 'in', dpi = 600)
```

#Running GLMM on Disease Status between treatment group over time
We found a significant interaction with change in disease status between time point, when we take out the interaction effect we will get different results but that might not matter if we want to look at ONLY Date and not treatment
***
First we are going to try and run Generalized Linear Mixed Models using `glmTMBB` is less finacky than glmer. 
We are going to look at the interaction between *Treatment* and *Date* with *DiseaseStatus* as the response variable and *CattleTag* as the random effect we will duse the *StatusData* dataframe. The family will be `binomial` since we have two response answers. 

***Getting an error with my random effect***
online it says : when a random-effect variance is estimated to be zero, or random-effect terms are estimated to be perfectly correlated ("singular fit": often caused by having too few levels of the random-effect grouping variable)

-removed random effect and it worked but then when i changed family to gaussian and left in the random effect that worked also

Also changed with family from  `binomial` to `gaussian` and that helped the model too

```{r}
StatusGLMM <- glmmTMB(DiseaseStatus ~ Treatment*Date + (1|CattleTag), data = StatusData, family ="gaussian")
summary(StatusGLMM) $r.squared
Anova(StatusGLMM, type=3)

```
#Running PostHoc test to understand significant interactions
Trying to figure out effect size using emmeans but not sure how to 

```{r}
EMM = emmeans(StatusGLMM, ~Date|Treatment)
EMM
eff_size(EMM, sigma(StatusGLMM), edf=df.residual(StatusGLMM))

confint(contrast(EMM, "pairwise", type = "Date"))

effectsize= standardize_parameters(StatusGLMM, exp = TRUE)

```


```{r}
emmeans(StatusGLMM, ~Date*Treatment)
```
# Running Post Hoc test for the GLMM above; time was significant in interactions
***
This didn't work until I made date a factor 

```{r}
Status.Post <-glht(StatusGLMM, mcp(Date='Tukey'))
summary(Status.Post)
lsmeans(StatusGLMM, pairwise~Date)
```


```{r}

ActiveProportionData$Date <- factor(ActiveProportionData$Date, labels=c(c("Sept 22","Sept 30", "Oct 10","Oct 22", "Nov 3", "Dec 1")))
#ActiveProportionData$ExperimentTime <- factor(ActiveProportionData$ExperimentTime, labels=c(c("Pre Treatment", "During Treatment", "Post Treatment")))
ActiveProportionData$ExperimentTime <- factor(ActiveProportionData$ExperimentTime, levels =c("PreTreatment", "DuringTreatment", "PostTreatment"), labels=c("Pre Treatment", "During Treatment", "Post Treatment"))
ActiveProportionData
```

#Visualizing that active/inactive data through proportions  
```{r}
ActiveProportionBarGraph <- ggplot(ActiveProportionData, aes(Treatment,Proportion, fill= Status, group=Treatment)) +
  geom_bar(position="stack",stat="identity") +
  facet_grid(~Date + ExperimentTime, labeller= labeller(label_parsed))+
  scale_fill_manual(values=c('#882255','#44AA99'), name="SCTLD Status")+
  scale_x_discrete (labels=c("Nutrient \n Amended", "Control")) +
  #From colorblind palette
  #Scale_fill_manual fills the bars with specific colors
  

      labs(y="Proportion of Active Colonies") +
 #Change the x axis labels
  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"))
      
ActiveProportionBarGraph
ggsave("Figures/SuppFigure3_ActiveProportion.BarGraph.JPG", plot = ActiveProportionBarGraph, height = 6, width = 12, units = 'in', dpi = 600)
```


# Running a GLMM to see if there is any significance between lesions over treatment and date 
***
There are no extreme outliers 

```{r}
StatusData%>%
  group_by(TimePoint, Treatment) %>%
identify_outliers(Lesions)

```
#Visualizing the Lesion Data
***
Not really showing anything worth discussing in thesis 

```{r}
Lesion.Bxp <- ggboxplot(StatusData, x="Date",y="Lesions", fill="Treatment" ) 
 Lesion.Bxp
```

```{r}
LesionGLMM <- glmmTMB(Lesions ~ Treatment*Date + (1|CattleTag), data = StatusData, family ="gaussian")
summary(LesionGLMM)
Anova(LesionGLMM, type=3)
lsmeans(LesionGLMM, pairwise ~ Treatment*Date, adjust="tukey")
```
```{r}
emmip(LesionGLMM, Treatment~Date)
```


#ENVIRONMENTAL FACTORS

***
First we will check the *Temp* and *Light* data for normalicy using the *StatusData*
We want to first check if there are any differences between *Date* and *Treatment*
***
We can see the data is not normally distributed for either *Temp* or *Light*, *Rain* is also not normally distributed, need to remove grouping for rain

```{r}
StatusData %>%
 group_by(Treatment,Date) %>%
  shapiro_test(Temp)
 
```

```{r}
Temp.bxp <- ggboxplot(StatusData, x="Date", y="Temp", fill="Treatment")
Temp.bxp
```

#Running a GLMM on Temp data


*** 
GLMM says there is a significant interaction between *Treatment* and *Date* on *Temp*. However, when we run a pairwise test there is only significance across timepoints. 

```{r}
TempGLMM <- glmmTMB(Temp ~ Treatment*Date + (1|CattleTag), data = StatusData, family ="gaussian")
summary(TempGLMM) $r.squared
Anova(TempGLMM, type=3)
```

```{r}

lsmeans(TempGLMM, pairwise~Treatment)
```
***
No we are trying to see if we can add temp into the disease status GLMM as a predictor variable, we see that date is till significant only on disease status. Running the pairwise comparison gives us the 10-10-21: 10-22-21 date specificially. 
***

```{r}
Active.TempGLMM <- glmmTMB(DiseaseStatus ~ Temp + Rain + (1|CattleTag), data = StatusData, family ="gaussian")
summary(Active.TempGLMM) $r.squared
Anova(Active.TempGLMM, type=3)
```

#Rain Differences 
***


```{r}
RainGLMM <- glm(Rain ~ Date, data = StatusData, family ="gaussian")
summary(RainGLMM) $r.squared
Anova(RainGLMM, type=3)

```
```{r}
lsmeans(RainGLMM, pairwise~Date)
```
#rain Data to double check the above using `Kruskal-Wallis` on *Rain* over *Date*
Shows significance across almost all dates
```{r}
Rain.Kruskal <-StatusData %>% kruskal_test(Rain ~ Date)
Rain.Kruskal
StatusData %>% kruskal_effsize(Rain ~ Date)
```

```{r}
dunnTest(Rain~Date, data=StatusData, method="bh")
```


```{r}
StatusData$Date <- factor(StatusData$Date, labels=c(c("Sept 22","Sept 30", "Oct 10","Oct 22", "Nov 3", "Dec 1")))
TempData <- StatusData %>%
  group_by(Date) %>%
  summarize(Temp=mean(Temp), Proportion=mean(ActiveProportion))
TempData$Temp = ((TempData$Temp-32) * (5/9))
  
TempData

```
#Visualizng temperature on disease status
Using *StatusProportionData* only used for this graph and correlation test 

```{r}
StatusProportionTemp <- StatusProportionData %>% group_by(Date)
StatusProportionTemp$Date <- as.factor(StatusProportionTemp$Date)


StatusProportionTemp$Date <- factor(StatusProportionData$Date, labels=c(c("Sept 22","Sept 30", "Oct 10","Oct 22", "Nov 3", "Dec 1")))
StatusProportionTemp
```
```{r}
shapiro_test(StatusProportionTemp$Temp)
shapiro_test(StatusProportionTemp$StatusProportion)
```

 *** Effect of rain on disease status
```{r}
cor.test(StatusProportionTemp$StatusProportion, StatusProportionTemp$Rain, method="pearson")
```
***
Because we saw a significant decline in temperature and active disease status a Pearson’s correlation test was performed. As temperature decreased the proportion of SCTLD active colonies also significantly decreased (Pearson correlation: t4 = 2.88, p = 0.045, r = 0.82).  
```{r}
cor.test(StatusProportionTemp$StatusProportion, StatusProportionTemp$Temp, method="pearson")
```
```{r}
Active_Rain.LineGraph <- ggplot(data=StatusProportionTemp, aes(x=Temp, y=StatusProportion)) +
  geom_line(color='#CC6677') +
  geom_point(aes(shape=Date), size=4)+
  #Adds a shape to each point based on date,and then the size of the spaes
  labs(y= "Proportion of Active SCTLD Colonies", x="Temperature (°C)" )+


  scale_shape_manual(values=c(15,25,17,23,19,22))+
#Manually chose shape numbers to assign to points
  
  geom_hline(yintercept=0.5000000, linetype=4, color="black")+
  #Adding a horizontal line through graph and the specific y value
  geom_text(aes(x = 29.5, label="End Treatment", y=.48), colour="black")+
#Adding label underneath the line above, will put label and x and y value
  scale_x_reverse(breaks = c(26,27,28,29,30), labels=c("26°","27°","28°", "29°","30°")) +
  #this reverses the scale 
  #scale_x_continuous(breaks = c(26,27,28,29,30), labels=c("26°","27°","28°", "29°","30°")) + 
  #the above works if we werent reversing the scale

  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"),
        legend.key=element_blank())
```

```{r}
install.packages(microshades)
library(microshades)

```

```{r}
Color <- c(rev(microshades_palette("micro_cvd_purple", 5)), (color="mistyrose2"))
           #microshades_palette("micro_purple",1, lightest=TRUE))
```


```{r}

Active_Temp.LineGraph <- ggplot(data=StatusProportionTemp, aes(x=Temp, y=StatusProportion)) +
  geom_line(color='#CC6677') +
  geom_point(aes(color=Date ), size=4)+
  
 scale_color_manual(values=c(Color)) +
  #scale_fill(low="blue", high="red") +
  #Adds a shape to each point based on date,and then the size of the spaes
  labs(y= "Proportion of Active SCTLD Colonies", x="Temperature (°C)" )+


 # scale_shape_manual(values=c(15,25,17,23,19,22))+
#Manually chose shape numbers to assign to points
  
  geom_hline(yintercept=0.5000000, linetype=4, color="black")+
  #Adding a horizontal line through graph and the specific y value
  geom_text(aes(x = 29.5, label="End Treatment", y=.48), colour="black")+
#Adding label underneath the line above, will put label and x and y value
  scale_x_reverse(breaks = c(30,29,28,27,26), labels=c("30°","29°","28°", "27°","26°")) +
  #this reverses the scale 
  #scale_x_continuous(breaks = c(26,27,28,29,30), labels=c("26°","27°","28°", "29°","30°")) + 
  #the above works if we werent reversing the scale

  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"),
        legend.key=element_blank())

      
Active_Temp.LineGraph
ggsave("Figures/Active_Temp.LineGraph.JPG", plot = Active_Temp.LineGraph, height = 5, width = 9, units = 'in', dpi = 600)
```
 #Running other environmental factors
 ***
 I tried to run other environmental factors but there weren't any other comparisons as above
 *** 
 Stick to 
 
```{r}
StatusData$Date <- factor(StatusData$Date, labels=c(c("Sept 22","Sept 30", "Oct 10","Oct 22", "Nov 3", "Dec 1")))
DIN.Data <- StatusData %>%
  group_by(Date) %>%
  summarize(DIN=mean(DIN), Proportion=mean(ActiveProportion), Rain = mean(Rain), AreaProportion= mean(AreaProportion), Light=mean(Light) )

  
DIN.Data
```

```{r}
ggplot(data=DIN.Data, aes(x=DIN, y=Proportion)) +
  geom_line(color="Violetred") +
  geom_point(aes(shape=Date), size=4)

```

```{r}
ActiveProportionData2 <- ActiveProportionData %>% filter(Status=="Active")
```

```{r}
Proportion_Nutrients <- ActiveProportionData2 %>%
  group_by(Date) %>%
  summarize(proportion=mean(Proportion), NO23 = mean(NO23), NH4=mean(NH4))
```


```{r}
ggplot(Proportion_Nutrients, aes(x=Date, group=1)) +
geom_line(color="blue", (aes(y=proportion))) +

  geom_line(aes(y=NO23), color="red") +
  #geom_point(color="green", aes(y=NH4))+
  geom_line(color="green", aes(y=NH4))
```

#Differences Between Timepoints

To look at differences across time I added the timepoint to the cattltag number. Example 501 on timepoint 3 is 501+3. 
```{r}
AreaProportionData <- ModelRateData %>% group_by(CattleTag) %>%
  summarize(Rate=mean(Rate))
AreaProportionData
```
#No significant difference between colonies
Doesn't seem to be any signficant difference between the colonies 


```{r}
Colony.Kruskal <- ModelRateData %>% kruskal_test(Rate~CattleTag)
Colony.Kruskal

```

```{r}

 ColonyProportion<- glm(DiseaseStatus ~ CattleTagTimepoint*Date, data = ModelRateData, family ="binomial")

summary(ColonyProportion)

```
```{r}
lsmeans(ColonyProportion, pairwise~CattleTagTimepoint)
```
#Trying to look at the effect of diseasestatus on rate

# Testing if there is an effect of disease status on change of proportion of tissue 

****
Find there to be significant changes in tissue area proportion throughout time points 
#Tried running a glmm but giving me difference answers than the Permanova
```{r}
Proportion.StatusGLMM <- glmmTMB(AreaProportion ~ DiseaseStatus*Date + (1|CattleTag), data = StatusData, family ="gaussian")
Anova(Proportion.StatusGLMM, type=3)

```

```{r}
Proportion.Status.Post <-glht(Proportion.StatusGLMM, linfct=mcp(Date ='Tukey'))
summary(Proportion.Status.Post)
lsmeans(Proportion.StatusGLMM, pairwise ~ Date*DiseaseStatus)
```

#Also want to look at the effect of status on rate

```{r}
Rate.StatusGLMM <- glmmTMB(Rate ~ DiseaseStatus*Date + (1|CattleTag), data = StatusData, family ="gaussian")
Anova(Rate.StatusGLMM, type=3)

```

```{r}
ActiveProportionData <- filter(ModelRateData, DiseaseStatus == "Active")
ActiveProportionData
```

```{r}
DiseaseStatusANOVA <- anova_test(data=ModelRateData,  AreaProportion~DiseaseStatus*Date)
DiseaseStatusANOVA
```
```{r}
model <- lm(AreaProportion ~ DiseaseStatus*Date, data= ModelRateData)
ModelRateData %>%
  group_by(DiseaseStatus) %>%
  anova_test(AreaProportion ~ Date, error=model)
```

```{r}
ModelRateData %>%
  group_by(DiseaseStatus) %>%
  pairwise_t_test(AreaProportion ~ Date, p.adjust.method="bonferroni")
```

```{r}
ModelRateData %>%
  group_by(Date) %>%
  pairwise_t_test(AreaProportion ~ DiseaseStatus, p.adjust.method="bonferroni")
```

#Visualizing the difference betwen active and inactive colonies rate of tissue loss throughout time
Creating a line graph for *Rate per day* over *Date* between the two *treatments*
**Also trying to look at DIN.SRP ratios on Active Status but need to look more into that** 



```{r}
AreaProportionStatus.Summary <- summarySE(ModelRateData, measurevar = "AreaProportion", groupvars = c("Date", "DiseaseStatus"), na.rm=TRUE) 
                                          
AreaProportionStatus.Summary
```

```{r} 
 ActiveTissueChange.LineGraph <- ggplot(AreaProportionStatus.Summary,aes(Date,AreaProportion,color=DiseaseStatus,group=DiseaseStatus)) +
  geom_line(aes(color=DiseaseStatus),size=1.5)+
  geom_vline(xintercept= as.factor("2021-10-22"), linetype=4, color="black")+
    geom_hline(yintercept=0, color="black") +
  
  
#adds intercept lint if not number have to do as.factor or whatever it might be
  geom_text(aes(x=as.factor("2021-10-22"), label="End Treatment", y=-.11), colour="black",angle=90, vjust=1.5)+
  #adds label to intercept line
  scale_color_manual(values=c('#882255',"#44AA99"), name="SCTLD Status") +
#changes colors and legend title+
  geom_errorbar(aes(ymin=AreaProportion-se, ymax=AreaProportion+se, color=DiseaseStatus), size=1.5,width=0.5) +
  

  #adds colors to order of variables


  
  #scale_y_continuous(breaks=c(0,5,10,15),limits = c(0, 15), labels=c("0%","5%", "10%", "15%"))

 scale_x_discrete(limits=c("2021-09-30", "2021-10-10","2021-10-22", "2021-11-03", "2021-12-01"), labels=c("Sept 30", "Oct 10", "Oct 22", "Nov 3", "Dec 1")) +
      labs(y="Percent of Tissue Change") +
    
  scale_y_continuous(labels = scales::percent_format(),breaks=c(-.15,-.10,-.05,0,.05,.10,.15), limits = c(-.15, .05)) +
 #Change the x axis labels
  theme(panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
        panel.background = element_rect(fill = '#F5F5F5'),
        plot.title = element_text(size = 23, face="bold", margin = margin(10,0,10,0)),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(color="black"),
        text = element_text(size=17, color="black"), 
        axis.text.x=element_text(angle = 45, hjust = 1,size=12, color="black"),
        axis.text.y=element_text(size=14, color="black"),
        legend.key=element_blank())
      
ActiveTissueChange.LineGraph
ggsave("Figures/Figure8_ActiveTissueChange.LineGraph.JPG", plot = ActiveTissueChange.LineGraph, height = 5, width = 9, units = 'in', dpi = 600)

```





